Our Galactocentric Coordinate System Is Probably Wrong, and Not by a Little

### 2018-12-06 paper/ms.tex
Update paper

\title{Our Galactocentric Reference Frame Is Probably Wrong, and Not by a
Little}

@@

In an axisymmetric potential the 6D phase space of stars' positions and
velocities can be reduced to a 3D space of invariant actions. We show that if
one assumes the wrong galactocentric reference frame this can induce
artificial non-axisymmetries which manifest as time dependence in the computed
actions. However, the Milky Way is not axisymmetric in detail and so there are
no truly axisymmetric axes. We instead define the  ``optimally axisymmetric''
axes as the reference frame which minimizes the time dependence of the
computed actions. Standard methods for determining the galactocentric
reference frame typically involve finding the median height of stars in the
disk. Using cosmological zoom-in simulations from FIRE-2, we show that the
optimally axisymmetric reference frame can differ from standard methods of
calculating the galactocentric reference frame by nearly a scale height. This
work calls into question the use of actions as dynamical invariants unless
these reference frame issues can be resolved.

% However, it is not expected that actions are invariant in detail for the
% Milky Way. This in-invariance can come about from assuming the wrong
% Galactic potential, but it can also come about due to true
% non-axisymmetries. We show that actions can gain artificial time-dependence
% also by assuming the wrong reference frame. We argue that standard methods
% for determining the galactocentric reference frame are not gu

% it's spiral arms, bar(s), and gas clouds are all non-axisymmetric features
% that should induce long-term evolution in action space. However, these same
% features will cause a star's true orbit to deviate slightly from its orbit
% in an axisymmetric potential. If an observer assumes an axisymmetric
% potential when computing actions, this will induce an artificial short-term
% phase dependence of the actions, which we interpret as systematic error.
% Using the potential of cosmological zoom-in simulations of Milky Way mass
% galaxies from the FIRE collaboration (the Latte suite), we investigate the
% magnitude of this systematic error. We find XXX. Something something lower
% limit.

@@

Underlying all of these discoveries is the assumption of a galactocentric
reference frame \citep{2008gady.book.....B}. In angular
coordinates, the center of the galaxy is taken to be the location of
Sgr~A\textsuperscript{*} \citep[e.g.][]{2004ApJ...616..872R}. From stellar
motions near Sgr~A\textsuperscript{*}, the distance to the center is taken to
be $\sim 8.1\text{--}8.3\,\kpc$
\citep{2009ApJ...692.1075G,2018A&A...615L..15G}. The vertical height is found
by identifying where the stellar positions and velocities reach a maximum
(effectively the median height of all disk stars), and is usually taken to be
$\sim 25\,\pc$ \citep{2001ApJ...553..184C}, with more recent measurements from
{\em Gaia} DR2 placing it at $\sim 20\,\pc$ \citep{2019MNRAS.482.1417B}.

These measurements of the galactocentric reference frame rely on the
assumption of an axisymmetric Milky Way. Under this assumption, these
estimates would provide the correct parameters for the galactocentric
reference frame up to measurement error. However, the Milky Way is not
axisymmetric in detail. Spiral arms, bar(s), and infalling satellite galaxies
such as Sagittarius and the Large Magellanic Cloud all induce
non-axisymmetries.

Since the Milky Way is not truly axisymmetric, it is not clear the best way to
define the galactocentric reference frame. In this work, we consider one
possible physically motivated approach to defining the galactocentric
reference frame, and show that it can differ from conventional methods of
frame measurements.

In a slowly-evolving axisymmetric potential, the actions of stars' orbits are
conserved quantities \citep{2008gady.book.....B,2014RvMP...86....1S}. In other
words, if we could measure the actions of a star in such a potential, we would
measure the same actions over the course of the orbit. The non-axisymmetries
of a more realistic potential induce time evolution in the {\em measured}
actions --- the true orbital invariants (if they exist) are invariant by
definition. Through simple tests, we will see that even in a truly
axisymmetric potential, if the wrong reference frame is assumed then there
will be additional time evolution in the measured actions.

@@

\section{Reference Frame Gone Wrong} \label{sec:ref_frame}
We show and quantify the time evolution of measured actions that comes about
from inaccuracies in the galactocentric reference frame.

@@

Now suppose this observer makes a measurement of the true orbit at the orange
point or the red point. Then the orange and red lines correspond to the orbits
that the observer would compute for each point. In action space, this would
correspond to a different value of $J_z$ for the orange and red points. In
this way, assuming the wrong reference frame induces time dependence in the
actions {\em computed by an observer}.

@@

\subsection{Quantification} \label{ssec:quant}
We now turn to classical orbital integration in a commonly assumed Milky Way
potential to quantify the time variation of computed actions for different
reference frame errors.

@@

First, we illustrate the time-dependence of one possible inaccurate reference
frame. We begin by taking an orbit with initial position $(8, 0, 0)\,\kpc$ and
initial velocity $(0, -190, 50)\,\kms$ --- this corresponds to an orbit with
actions $(J_r, L_z, J_z) = (32.5, -1520, 23.0)\,\actunit$. The orbit is
plotted in \Gus{Appendix x}. We then suppose that an observer makes
observations of this orbit in a reference frame that is incorrect in radius by
$200\,\pc$ and height by $100\,\pc$ --- we subtract the vector $(200, 0,
100)\,\pc$ from each position in the orbit. This corresponds to an observer
physically located at the position $(8, 0, 0)\,\kpc$ but erroneously thinking
they are located at $(8.2, 0, 0.1)\,\kpc$.

Using the observations of the orbit that this erroneous observer makes every
$\Myr$, we perform the same orbit integration procedure as before to compute
the actions our observer would make in the incorrect reference frame over the
duration of the orbit. The erroneous computed actions are shown for the first
$\Gyr$ of the orbit in Figure~\ref{fig:one_orbit_wrong_ref}. Occasionally the
numerical scheme fails and very large actions are reported by \texttt{gala}
--- we perform a $5\sigma$ clip on each action to exclude such orbits, but
this only excludes a total of $5$ orbits out of the $1000$ considered for
Figure~\ref{fig:one_orbit_wrong_ref}. Some numerical artifacts remain, but the
vast majority of orbits are computed properly.

@@

\section{Optimally Axisymmetric Frame} \label{sec:oa_frame}
An analytic, axisymmetric potential with a known reference frame is a far too
simple description of the Milky Way. We thus turn to cosmological zoom-in
simulations as approximations to a ``true'' Milky Way potential. We will
perform a few simulations of small clusters embedded in the potential of a
frozen $\z=0$ snapshot of one such zoom-in. We take our fiducial reference
frame from \citet{2018arXiv180610564S}.

@@

Once we perform this simulation, we calculate the actions of each star in the
cluster assuming the best-fit smooth, axisymmetric potential, as we do
observationally for the Milky Way. We refer to this space as the observational
action space. We will find that there is significant time evolution in
observational action space. Attributing some of this evolution to our fiducial
reference frame being incorrect, we minimize the time-dispersion in action
space

\subsection{Cosmological Zoom-in} \label{ssec:cosmozoom}
Cosmological zoom-ins allow one to simulate a selected region at high
resolution embedded in a low-resolution cosmological background
\citep[e.g.][]{1993ApJ...412..455K,2014MNRAS.437.1894O}. Here we use a
zoom-in of a Milky Way-mass galaxy from the FIRE-2 collaboration.

First, a dark matter only simulation was run with $\Lambda$CDM cosmology
parameterized by $\Omega_m = 0.272$, $\Omega_{\Lambda} = 0.728$, $\Omega_b =
0.0455$, $h = 0.702$, $\sigma_8 = 0.807$, and $n_s = 0.961$, consistent with
current constraints \citep{2018arXiv180706209P}. Halos are selected at $z=0$
based solely on their mass ($\sim 1\text{--}2 \times 10^{12} \Msun$).
Particles within $5 R_{200\text{m}}$ of each halo are traced back to $\z=99$
and the initial conditions are regenerated at higher resolution using MUSIC
\citep{2011MNRAS.415.2101H}. A total of $\sim 10$ halos were simulated at
higher resolution \citep{2018MNRAS.481.4133G}.

We restrict ourselves to the simulation \texttt{m12i} \citep[first introduced
in][]{2016ApJ...827L..23W}. For this work, this galaxy is the best-case
scenario. It does not contain a bar (like \texttt{m12m}) or companion galaxy,
and by eye is the closest to axisymmetric of the three galaxies presented in
\citet{2018arXiv180610564S} \citep{2018MNRAS.481.4133G}.

@@

To compute the forces from the frozen potential, we use the simple tree code
\texttt{pykdgrav}\footnote{\url{https://github.com/omgspace/pykdgrav},
commit: \texttt{fbd1ddc}}. We use an opening angle of $0.5$ and exclude stars
at distances $R>50\,\kpc$ in the fiducial reference frame.

We use the initial position and velocity of two star particles in
$\texttt{m12i}$. These two star particles are randomly chosen from all star
particles in the simulation with the following properties (in the fiducial
reference frame):
\begin{enumerate}
    \item age in the range $(0.25, 0.75)\,\Gyr$,
    \item cylindrical radius in the range $(7.7, 8.7)\,\kpc$,
    \item height (absolute value) less than $0.5\,\kpc$,
    \item $J_r$ and $J_z$ in the range:
        \subitem $(10, 20)\,\actunit$ for the first star,
        \subitem $(40, 60)\,\actunit$ for the second star.
\end{enumerate}
The ids we choose are \texttt{23693026} for the first star and
\texttt{17012804} for the second star. We refrain from calling these ``low''
or ``high'' action stars since the actions are computed in a presumably
incorrect reference frame.

@@

\subsection{Frame Optimization} \label{ssec:frame_opt}
Figure~\Gus{something} shows the action computed at each time step in the
fiducial reference frame. Clearly there is significant time evolution.

Even though we change the center, we do not recalculate the center velocity.
This is largely irrelevant, as the recalculated center velocity differs from
the fiducial center velocity by $<0.1\,\kms$. In recentering the galaxy, the
package \texttt{gizmo\_analysis} will subtract a slightly different value for
the Hubble flow from each particle's velocity to convert to a physical
reference frame. The error this introduces is $\sim 0.02\,\kms$, which we also
ignore.

>>>>

\title{Our Galactocentric Coordinate System Is Probably Wrong, and Not by a
Little}

@@

In a completely axisymmetric potential, the notion of a global Galactic
midplane is well defined. However, since the Milky Way is not completely
axisymmetric, the global Galactic midplane is to some level of precision not
well defined. The local Galactic midplane has been found by determining the
height at which the number counts of stars reaches a maximum. This local
midplane is then used as the global midplane, an assumption which would be
justified in a completely axisymmetric potential. Assuming a global Galactic
midplane exists, we show in three cosmological zoom-in simulations from the
FIRE collaboration that local determinations of the midplane can vary by
nearly a thin disk scale height. Using standard orbit integration techniques,
we show that such an inaccuracy in the midplane can induce artifical time
evolution in the vertical action that makes it virtually unusable for
individual stars in the disk. The situation is less dire for angular momentum

% Local determinations of the Galactic midplane rely on finding the height at
% which the number count of stars reaches a maximum. This is not guaranteed to
% be

% In an axisymmetric, adiabatic potential, the 6D phase space of stars' positions and velocities can be reduced to a 3D space of invariant actions. We show that if one assumes the wrong galactocentric coordinate system this can induce significant artificial time dependence in the computed actions.

% In an axisymmetric potential the 6D phase space of stars' positions and
% velocities can be reduced to a 3D space of invariant actions. We show that if
% one assumes the wrong galactocentric coordinate system this can induce
% artificial non-axisymmetries which manifest as time dependence in the computed
% actions. However, the Milky Way is not axisymmetric in detail and so there are
% no truly axisymmetric axes. We instead define the  ``optimally axisymmetric''
% axes as the coordinate system which minimizes the time dependence of the
% computed actions. Standard methods for determining the galactocentric
% coordinate system typically involve finding the median height of stars in the
% disk. Using cosmological zoom-in simulations from FIRE-2, we show that the
% optimally axisymmetric coordinate system can differ from standard methods of
% calculating the galactocentric coordinate system by nearly a scale height.
% This work calls into question the use of actions as dynamical invariants
% unless these coordinate system issues can be resolved.

@@

Underlying all of these discoveries is the assumption of a galactocentric
coordinate system \citep{2008gady.book.....B}. In angular
coordinates, the center of the galaxy is taken to be the location of
Sgr~A\textsuperscript{*} \citep[e.g.][]{2004ApJ...616..872R}. From stellar
motions near Sgr~A\textsuperscript{*}, the distance to the center is taken to
be $\sim 8.1\text{--}8.3\,\kpc$
\citep{2009ApJ...692.1075G,2018A&A...615L..15G}. The vertical height is found
by identifying where the stellar positions and velocities reach a maximum
(effectively the median height of all disk stars), and is usually taken to be
$\sim 25\,\pc$ \citep{2001ApJ...553..184C}, with more recent measurements from
{\em Gaia} DR2 placing it at $\sim 20\,\pc$ \citep{2019MNRAS.482.1417B}.

These measurements of the galactocentric coordinate system rely on the
assumption of an axisymmetric Milky Way. Under this assumption, these
estimates would provide the correct parameters for the galactocentric
coordinate system up to measurement error. However, the Milky Way is not
axisymmetric in detail. Spiral arms, bar(s), and infalling satellite galaxies
such as Sagittarius and the Large Magellanic Cloud all induce
non-axisymmetries.

Since the Milky Way is not truly axisymmetric, it is not clear the best way to
define the galactocentric coordinate system. In this work, we consider one
possible physically motivated approach to defining the galactocentric
coordinate system, and show that it can differ from conventional methods of
frame measurements.

In a slowly-evolving axisymmetric potential, the actions of stars' orbits are
conserved quantities \citep{2008gady.book.....B,2014RvMP...86....1S}. In other
words, if we could measure the actions of a star in such a potential, we would
measure the same actions over the course of the orbit. The non-axisymmetries
of a more realistic potential induce time evolution in the {\em measured}
actions --- the true orbital invariants (if they exist) are invariant by
definition. Through simple tests, we will see that even in a truly
axisymmetric potential, if the wrong coordinate system is assumed then there
will be additional time evolution in the measured actions.

@@

\section{Local Midplane in FIRE} \label{sec:local_fire}
\subsection{Description of Simulation} \label{ssec:cosmozoom}
Here we briefly describe the cosmological zoom-ins used in this work.
Cosmological zoom-ins allow one to simulate a selected region at high
resolution embedded in a low-resolution cosmological background
\citep[e.g.][]{1993ApJ...412..455K,2014MNRAS.437.1894O}. The FIRE
collaboration has simulated a number of Milky Way-mass zoom-ins as part of the
{\em Latte} suite of FIRE-2 simulations
\citep{2016ApJ...827L..23W,2018MNRAS.481.4133G}. We use the three zoom-ins
considered in \citet{2018arXiv180610564S} --- \texttt{m12i}, \texttt{m12f},
\texttt{m12m}, since they have associated mock {\em Gaia} DR2 catalogues and
the $z=0$ snapshots are publicly available.

First, a dark matter only simulation was run with $\Lambda$CDM cosmology
parameterized by $\Omega_m = 0.272$, $\Omega_{\Lambda} = 0.728$, $\Omega_b =
0.0455$, $h = 0.702$, $\sigma_8 = 0.807$, and $n_s = 0.961$, consistent with
current constraints \citep{2018arXiv180706209P}. Halos are selected at $z=0$
based solely on their mass ($\sim 1\text{--}2 \times 10^{12} \Msun$).
Particles within $5 R_{200\text{m}}$ of each halo are traced back to $\z=99$
and the initial conditions are regenerated at higher resolution using MUSIC
\citep{2011MNRAS.415.2101H}.

Simulations were run using
\texttt{GIZMO}\footnote{\url{http://www.tapir.caltech.edu/~phopkins/Site/GIZMO.html}}
\citep{2015MNRAS.450...53H}. Hydrodynamics is solved using the meshless
finite-mass method and gravity is solved using a modified version of the
Tree-PM solver of \texttt{GADGET-3}, using fully adaptive and fully
conservative gravitational force softening for gas.

% We restrict ourselves to the simulation \texttt{m12i} \citep[first introduced
% in][]{2016ApJ...827L..23W}. For this work, this galaxy is the best-case
% scenario. It does not contain a bar (like \texttt{m12m}) or companion galaxy,
% and by eye is the closest to axisymmetric of the three galaxies presented in
% \citet{2018arXiv180610564S} \citep{2018MNRAS.481.4133G}.

\subsection{Local Midplane} \label{sec:local_midplane}
We

\section{Coordinate System Gone Wrong} \label{sec:ref_frame}
We show and quantify the time evolution of measured actions that comes about
from inaccuracies in the galactocentric coordinate system.

@@

Now suppose this observer makes a measurement of the true orbit at the orange
point or the red point. Then the orange and red lines correspond to the orbits
that the observer would compute for each point. In action space, this would
correspond to a different value of $J_z$ for the orange and red points. In
this way, assuming the wrong coordinate system induces time dependence in the
actions {\em computed by an observer}.

@@

\subsection{Quantification} \label{ssec:quant}
We now turn to classical orbital integration in a commonly assumed Milky Way
potential to quantify the time variation of computed actions for different
coordinate system errors.

@@

First, we illustrate the time-dependence of one possible inaccurate reference
frame. We begin by taking an orbit with initial position $(8, 0, 0)\,\kpc$ and
initial velocity $(0, -190, 50)\,\kms$ --- this corresponds to an orbit with
actions $(J_r, L_z, J_z) = (32.5, -1520, 23.0)\,\actunit$. The orbit is
plotted in \Gus{Appendix x}. We then suppose that an observer makes
observations of this orbit in a coordinate system that is incorrect in radius by
$200\,\pc$ and height by $100\,\pc$ --- we subtract the vector $(200, 0,
100)\,\pc$ from each position in the orbit. This corresponds to an observer
physically located at the position $(8, 0, 0)\,\kpc$ but erroneously thinking
they are located at $(8.2, 0, 0.1)\,\kpc$.

Using the observations of the orbit that this erroneous observer makes every
$\Myr$, we perform the same orbit integration procedure as before to compute
the actions our observer would make in the incorrect coordinate system over the
duration of the orbit. The erroneous computed actions are shown for the first
$\Gyr$ of the orbit in Figure~\ref{fig:one_orbit_wrong_ref}. Occasionally the
numerical scheme fails and very large actions are reported by \texttt{gala}
--- we perform a $5\sigma$ clip on each action to exclude such orbits, but
this only excludes a total of $5$ orbits out of the $1000$ considered for
Figure~\ref{fig:one_orbit_wrong_ref}. Some numerical artifacts remain, but the
vast majority of orbits are computed properly.

@@

\section{Optimally Axisymmetric Frame} \label{sec:oa_frame}
An analytic, axisymmetric potential with a known coordinate system is a far
too simple description of the Milky Way. We thus turn to cosmological zoom-in
simulations as approximations to a ``true'' Milky Way potential. We will
perform a few simulations of small clusters embedded in the potential of a
frozen $\z=0$ snapshot of one such zoom-in. We take our fiducial reference
frame from \citet{2018arXiv180610564S}.

@@

Once we perform this simulation, we calculate the actions of each star in the
cluster assuming the best-fit smooth, axisymmetric potential, as we do
observationally for the Milky Way. We refer to this space as the observational
action space. We will find that there is significant time evolution in
observational action space. Attributing some of this evolution to our fiducial
coordinate system being incorrect, we minimize the time-dispersion in action
space

@@

To compute the forces from the frozen potential, we use the simple tree code
\texttt{pykdgrav}\footnote{\url{https://github.com/omgspace/pykdgrav},
commit: \texttt{fbd1ddc}}. We use an opening angle of $0.5$ and exclude stars
at distances $R>50\,\kpc$ in the fiducial coordinate system.

We use the initial position and velocity of two star particles in
$\texttt{m12i}$. These two star particles are randomly chosen from all star
particles in the simulation with the following properties (in the fiducial
coordinate system):
\begin{enumerate}
    \item age in the range $(0.25, 0.75)\,\Gyr$,
    \item cylindrical radius in the range $(7.7, 8.7)\,\kpc$,
    \item height (absolute value) less than $0.5\,\kpc$,
    \item $J_r$ and $J_z$ in the range:
        \subitem $(10, 20)\,\actunit$ for the first star,
        \subitem $(40, 60)\,\actunit$ for the second star.
\end{enumerate}
The ids we choose are \texttt{23693026} for the first star and
\texttt{17012804} for the second star. We refrain from calling these ``low''
or ``high'' action stars since the actions are computed in a presumably
incorrect coordinate system.

@@

\subsection{Frame Optimization} \label{ssec:frame_opt}
Figure~\Gus{something} shows the action computed at each time step in the
fiducial coordinate system. Clearly there is significant time evolution.

Even though we change the center, we do not recalculate the center velocity.
This is largely irrelevant, as the recalculated center velocity differs from
the fiducial center velocity by $<0.1\,\kms$. In recentering the galaxy, the
package \texttt{gizmo\_analysis} will subtract a slightly different value for
the Hubble flow from each particle's velocity to convert to a physical
coordinate system. The error this introduces is $\sim 0.02\,\kms$, which we
also ignore.
